The state learner \\ {-- a super learner for right-censored data}

### 2024-05-03 tex/jrss-version/state-learner-jrss-style.tex
answered TAG's comments for sections 8-10

Using equation~(\ref{eq:surv-def}) we may write
\( \theta_{{\tau}}(Q) = \tilde{\Psi}_{t}^0(\Lambda_1, \Lambda_2, \mu) \), where
\begin{equation}
  \label{eq:1}    
  \begin{split}
  \tilde{\Psi}_{t}^0(\Lambda_1, \Lambda_2, \mu) & =
  \int_{\mathcal{W}} 
  \int_0^{\tau}
  e^{-\Lambda_1(s- \mid w, 1)-\Lambda_2(s- \mid w, 1)}  \Lambda_1(\diff s \mid
  w, 1)
  \mu(\diff w)
  \\
  &  \quad
  -\int_{\mathcal{W}} 
  \int_0^{\tau}
  e^{-\Lambda_1(s- \mid w, 0)-\Lambda_2(s- \mid w, 0)}  \Lambda_1(\diff s \mid w, 0)
  \mu(\diff w).
  \end{split}
\end{equation}
The efficient influence function for the parameter $\tilde{\Psi}_{\tau}$ depends
on the set \( (\Lambda_1, \Lambda_2, \Gamma, \pi) \) of nuisance parameters.
We define
\begin{align*}
  \omega_a(A,W; \pi)
  &=  \frac{% (-1)^{a+1}
    \1{\{A=a\}}}{\pi(W)^{a}(1-\pi(W))^{1-a}},
  \\
  g(t, A, W; \Lambda_1, \Lambda_2)
  & = \int_0^{t}
    e^{-\Lambda_1(s- \mid W, A)-\Lambda_2(s- \mid W, A)}  \Lambda_1(\diff s \mid
    W, A),
  \\  
  M_j(\diff t \mid A, W;  \Lambda_j  )
  & = N_j(\diff t) -
    \1{\{\tilde{T} \geq t\}} \Lambda_j(\diff t \mid W, A),
    \quad j \in \{1,2\},
  \intertext{and}
  M(\diff t \mid A, W;  \Lambda_1, \Lambda_2  )
  & = M_1(\diff t \mid A, W;  \Lambda_1  ) +
    M_2(\diff t \mid A, W;  \Lambda_2  ).
\end{align*}
The efficient influence function can now be written as
\citep{van2003unified,jewell2007non,rytgaard2022targeted},
\begin{equation*}
  \psi_{\tau}(O; \Lambda_1, \Lambda_2, \Gamma, \pi)
  = \psi_{\tau}^1(O; \Lambda_1, \Lambda_2, \Gamma, \pi)
  - \psi_{\tau}^0(O; \Lambda_1, \Lambda_2, \Gamma, \pi)
  -\tilde{\Psi}_{t}^0(\Lambda_1, \Lambda_2, \mu),
\end{equation*}
where
\begin{equation}
  \label{eq:5}
  \begin{split}
    & \psi_{\tau}^a(O; \Lambda_1, \Lambda_2, \Gamma, \pi)
    \\
    & =
      \omega_a(A,W; \pi)
      \int_0^{\tau} e^{\Gamma(t- \mid A, W)}   
      M_1(\diff t \mid A, W; \Lambda_1)
    \\
    & \quad
      -
      \omega_a(A,W; \pi)
      g(\tau, A, W; \Lambda_1, \Lambda_2)
      \int_0^{\tau}
      e^{[\Gamma+\Lambda_1 + \Lambda_2](t- \mid A, W)}
      M(\diff t \mid A, W; \Lambda_1, \Lambda_2)
    \\
    & \quad
      +
      \omega_a(A,W; \pi)      
      \int_0^{\tau}
      g(t, A, W; \Lambda_1, \Lambda_2)
      e^{[\Gamma+\Lambda_1 + \Lambda_2](t- \mid A, W)}
      M(\diff t \mid A, W; \Lambda_1, \Lambda_2)
    \\
    & \quad + g(\tau, a, W; \Lambda_1, \Lambda_2).
  \end{split}
\end{equation}

Equations~(\ref{eq:1}) and~(\ref{eq:5}) allow us to construct a one-step
estimator by using the definition given in equation~(\ref{eq:one-step-def}),
which gives the estimator
\begin{equation}
  \label{eq:one-step-comp-ate}
  \begin{split}
    \hat{\Psi}_{t,\text{OS}}(\data_n)
    = &
        \tilde{\Psi}_{t}^0(\hat{a}_1(\data_n), \hat{a}_2(\data_n),
        \hat{\mu}_n(\data_n))
    \\
      &
        +
        \empmeas{[\psi_{\tau}(\blank; \hat{a}_{1,n}(\data_n), \hat{a}_{2,n}(\data_n),
        \hat{b}_n(\data_n), \hat{\pi}_n(\data_n))]}
    \\
    = &
        \empmeas{[\psi_{\tau}^1(\blank; \hat{a}_{1,n}(\data_n), \hat{a}_{2,n}(\data_n),
        \hat{b}_n(\data_n), \hat{\pi}_n(\data_n))]}
    \\
      &
        - \empmeas{[\psi_{\tau}^0(\blank; \hat{a}_{1,n}(\data_n), \hat{a}_{2,n}(\data_n),
        \hat{b}_n(\data_n), \hat{\pi}_n(\data_n))]}.
  \end{split}
\end{equation}

@@

In this section we report results from two settings of our simulation
studies, both settings are without competing risks. In the next
section, we consider a setting with competing risks.
% ANDERS: kan du skrive her hvad forskellen er mellem og formaalet er
% med de to simulationsstudierscenarier?
% i den ene sammenligner vi med IPCW i den anden med westling, ik?
% Maaske en figur (flowchart) eller tabel kunne hjaelpe laeseren at faa overblik.
% Maaske separation i to subsections kunne hjaelpe yderliger

@@

We compare the state learner to two IPCW based super learners: The
first superlearner, called IPCW(Cox), uses a Cox model with main
effects to estimate the censoring probabilities, while the second
super learner, called IPCW(KM), uses the Kaplan-Meier estimator to
estimate the censoring probabilities. The Cox model for the censoring
distribution is correctly specified in both simulation settings while
the Kaplan Meier estimator only estimates the censoring model
correctly in the simulation setting where censoring is
independent. Both IPCW super learners are fitted using the
\texttt{R}-package \texttt{riskRegression}
\citep{Gerds_Ohlendorff_Ozenne_2023}.
% ANDERS: det er uklar for mig hvordan
% de super learners valger/kombinerer de learners der ingaar i library: er det Brier score? over time?
Each super learner provides a learner for the cumulative hazard
function for the outcome of interest. From the cumulative hazard
function a risk prediction model can be obtained (c.f.,
equation~(\ref{eq:surv-def}) with $\Lambda_2 = 0$). We measure the
performance of each super learner by calculating the index of
prediction accuracy (IPA) \citep{kattan2018index} at a fixed time
horizon (36 months) for the risk prediction model provided by the
super learner. The IPA is 1 minus the ratio between the model's Brier
score and the null model's Brier score, where the null model is the
model that does not use any covariate information. The IPA is
approximated using a large (\( n = 20,000 \)) independent data set of
uncensored data. As a benchmark we calculate the performance of the
risk prediction model chosen by the oracle selector, which uses the
large data set of uncensored event times to select the learner with
the highest IPA.

@@

\begin{figure}
  \centering %
  \includeFigCond[width=1\linewidth]{experiment-fig-sl-ipcw.pdf}
  \caption[]{For the risk prediction models provided by each of the super
    learners, the IPA is plotted against sample size. The results are averages across 
    1000 simulated data sets and the error bars are used to quantify the Monte Carlo
    uncertainty.
    % ANDERS: er det korrekt at det er averages across 1000 data sets? 
  }
  \label{fig:ipcw-fail}
\end{figure}

@@

The results are shown in Figure~\ref{fig:zelefski}. We see that for
most sample sizes, the state learner selected prediction models for
both censoring and outcome which have similar or higher IPA compared
to the prediction models selected by \texttt{survSL}.
% ANDERS: it is a bit lazy to write "for most sample sizes" :) it is
% slightly annoying that the y-axis scale changes in the 4
% figures. would that be better with Brier instead of IPA?

\begin{figure}
  \centering %
  \includeFigCond[width=1\linewidth]{experiment-fig-sl-survSL.pdf}
  \caption[]{For the risk prediction models of both the outcome and the
    censoring probability provided by each of the super learners, the IPA is
    plotted against sample size. The results are averages across 1000 repetitions and
    the error bars are used to quantify the Monte Carlo uncertainty.}
  % ANDERS: would be good to know how how many trainings set individuals were right censored before
  % the evaluation time point of the IPA. maybe of interest how you chose the evaluation time point.
  \label{fig:zelefski}
\end{figure}

@@

% ANDERS: har brug for en gentle intro saetning som fortaeller hvad der kommer: in this section we re-analyse the ...
%         Det skal ogsaa vaere klart i starten af sektion af der kommer noget om average treatment effect.
The prostate cancer data analyzed by \cite{kattan2000pretreatment},
which we introduced in Section~\ref{sec:numer-exper}, include a
competing event in the form of death without tumor recurrence. For
illustration, we fit the state learner to the original data set
consisting of 1,042 patients. We consider death without tumor
recurrence and recurrence of tumor as competing risks. We include the
fives learner \texttt{KM}, \texttt{Cox strata CS}, \texttt{Lasso},
\texttt{Elastic}, and \texttt{RF} which are described in
Table~\ref{tab:zel-library}. We use the same library of learners to
learn \( \Lambda_1 \), \( \Lambda_2 \), and $\Gamma$. In this case,
\( \Lambda_1 \) denotes the cause-specific cumulative hazard function
of tumor recurrence, and \( \Lambda_2 \) denotes the cause-specific
cumulative hazard function of death without tumor recurrence.

@@

As an illustration, we use the learners of the three cumulative hazard functions
selected by the state learner and the estimator defined in
Section~\ref{sec:cause-spec-aver}, equation~(\ref{eq:one-step-comp-ate}), to
estimate the average treatment effect of hormone therapy on risk of tumor
recurrence and death. The propensity score is estimated with a lasso model that
includes all levels of interaction. The results are shown in
Figure~\ref{fig:zelefski-real-target} for 6 month intervals after baseline with
point wise 95\% confidence intervals. We see that hormone therapy appears to
decrease the risk of tumor recurrence and increase the risk of death without
tumor recurrence, but that none of the estimated effects are statistically
significant from zero.

@@

We have proposed a new super learner that can be used with
right-censored data and competing events. Compared to existing
IPCW-based methods, the advantage of the state learner is that it does
not depend on a pre-specified estimator of the censoring distribution,
but selects one automatically based on a library of learners for the
censoring distribution. Furthermore, the state learner neither
requires evaluation of the density of \( \Lambda_j \) nor assumes a
% ANDERS: "density of \Lambda_j" er uklar hvordan er denne density defineret?
(semi-)parametric form for $\Lambda_j$. In the remainder of this
section we point to some limitations of our proposal and discuss
avenues for further research.

A major advantage of the state learner is that the performance of each
combination of learners can be estimated without (semi)parametric
assumptions or additional nuisance parameters. The drawback with this
approach is that we are rarely interested in features of the observed
data distribution when the data are right-censored: we are only
interested in the uncensored event time distribution. The finite
sample oracle inequality in Corollary~\ref{cor:oracle-prop} concerns
the function \( F \), which is a feature of \( P \in \mathcal{P} \),
while what we are typically interested in is \( \Lambda_j \) or
\( S \), which are features of \( Q \in \mathcal{Q} \). We emphasise
that while the state learner provides us with estimates of
\( \Lambda_j \) and $\Gamma$ based on libraries \( \mathcal{A}_j \)
and \( \mathcal{B} \), the performance of these learners is not
assessed directly for their respective target parameters, but only
indirectly via the performance of \( F \).
% ANDERS: er du enig med min omformulering af den sidste saetning?
For settings without competing risks, our numerical studies suggest
that measuring the performance of \( F \) also leads to good
performance for estimation of \( S \). However, further research on
this issue, both numerical and theoretical, is warranted.
% ANDERS: ved ikke helt hvad du taenker man kan goer i fremtiden men synes egentig at den sidste saetning kunne godt slettes. :)

@@

Our proposed super learner can be implemented with a broad library of learners
and using existing software.
% For instance the \texttt{R}-package
% \texttt{riskRegression} \citep{Gerds_Ohlendorff_Ozenne_2023} and the \texttt{R}-package by westling
Furthermore, while
the library \( \mathcal{F}(\mathcal{A}_1,\mathcal{A}_2,\mathcal{B}) \) consists
of \( |\mathcal{A}_1||\mathcal{A}_2||\mathcal{B}| \) many learners, we only need to fit
\( |\mathcal{A}_1| +|\mathcal{A}_2| + |\mathcal{B}| \) many learners in each fold. To
evaluate the performance of each learner we need to perform
\( |\mathcal{A}_1||\mathcal{A}_2||\mathcal{B}| \) many operations to calculate the
integrated Brier score in each hold-out sample, one for each combination of the
fitted models, but these operations are often negligible compared to fitting the
models. Hence the state learner is essentially not more computationally demanding
than any procedure that uses super learning to learn $\Lambda_1$, $\Lambda_2$,
and $\Gamma$ separately. While our proposal is based on constructing the library
\( \mathcal{F} \) from libraries for learning \( \Lambda_1 \), $\Lambda_2$, and
$\Gamma$, it would also be of interest to consider learners that estimate
\( F \) directly.
% ANDERS: how would one estimate F? using HAL? or RF? but such a learner would not provide risk prediction models or would it? if yes you could say it

>>>>

Using equation~(\ref{eq:surv-def}) we may write
\( \theta_{{\tau}}(Q) = \tilde{\Psi}_{t}^0(\Lambda_1, \Lambda_2, \mu) \), where
\begin{equation}
  \label{eq:1}    
  \begin{split}
  \tilde{\Psi}_{t}^0(\Lambda_1, \Lambda_2, \mu) & =
  \int_{\mathcal{W}} 
  \int_0^{\tau}
  e^{-\Lambda_1(s- \mid w, 1)-\Lambda_2(s- \mid w, 1)}  \Lambda_1(\diff s \mid
  w, 1)
  \mu(\diff w)
  \\
  &  \quad
  -\int_{\mathcal{W}} 
  \int_0^{\tau}
  e^{-\Lambda_1(s- \mid w, 0)-\Lambda_2(s- \mid w, 0)}  \Lambda_1(\diff s \mid w, 0)
  \mu(\diff w).
  \end{split}
\end{equation}
The efficient influence function for the parameter $\tilde{\Psi}_{\tau}$ depends
on the set \( (\Lambda_1, \Lambda_2, \Gamma, \pi) \) of nuisance parameters.
We define
\begin{align*}
  \omega_a(A,W; \pi)
  &=  \frac{% (-1)^{a+1}
    \1{\{A=a\}}}{\pi(W)^{a}(1-\pi(W))^{1-a}},
  \\
  g(t, A, W; \Lambda_1, \Lambda_2)
  & = \int_0^{t}
    e^{-\Lambda_1(s- \mid W, A)-\Lambda_2(s- \mid W, A)}  \Lambda_1(\diff s \mid
    W, A),
  \\  
  M_j(\diff t \mid A, W;  \Lambda_j  )
  & = N_j(\diff t) -
    \1{\{\tilde{T} \geq t\}} \Lambda_j(\diff t \mid W, A),
    \quad j \in \{1,2\},
  \intertext{and}
  M(\diff t \mid A, W;  \Lambda_1, \Lambda_2  )
  & = M_1(\diff t \mid A, W;  \Lambda_1  ) +
    M_2(\diff t \mid A, W;  \Lambda_2  ).
\end{align*}
The efficient influence function can now be written as
\citep{van2003unified,jewell2007non,rytgaard2022targeted},
\begin{equation*}
  \psi_{\tau}(O; \Lambda_1, \Lambda_2, \Gamma, \pi)
  = \psi_{\tau}^1(O; \Lambda_1, \Lambda_2, \Gamma, \pi)
  - \psi_{\tau}^0(O; \Lambda_1, \Lambda_2, \Gamma, \pi)
  -\tilde{\Psi}_{t}^0(\Lambda_1, \Lambda_2, \mu),
\end{equation*}
where
\begin{equation}
  \label{eq:5}
  \begin{split}
    & \psi_{\tau}^a(O; \Lambda_1, \Lambda_2, \Gamma, \pi)
    \\
    & =
      \omega_a(A,W; \pi)
      \int_0^{\tau} e^{\Gamma(t- \mid A, W)}   
      M_1(\diff t \mid A, W; \Lambda_1)
    \\
    & \quad
      -
      \omega_a(A,W; \pi)
      g(\tau, A, W; \Lambda_1, \Lambda_2)
      \int_0^{\tau}
      e^{[\Gamma+\Lambda_1 + \Lambda_2](t- \mid A, W)}
      M(\diff t \mid A, W; \Lambda_1, \Lambda_2)
    \\
    & \quad
      +
      \omega_a(A,W; \pi)      
      \int_0^{\tau}
      g(t, A, W; \Lambda_1, \Lambda_2)
      e^{[\Gamma+\Lambda_1 + \Lambda_2](t- \mid A, W)}
      M(\diff t \mid A, W; \Lambda_1, \Lambda_2)
    \\
    & \quad + g(\tau, a, W; \Lambda_1, \Lambda_2).
  \end{split}
\end{equation}
Equations~(\ref{eq:1}) and~(\ref{eq:5}) allow us to construct a one-step
estimator by using the definition given in equation~(\ref{eq:one-step-def}),
which gives the estimator
\begin{equation}
  \label{eq:one-step-comp-ate}
  \begin{split}
    \hat{\Psi}_{t,\text{OS}}(\data_n)
    = &
        \tilde{\Psi}_{t}^0(\hat{a}_1(\data_n), \hat{a}_2(\data_n),
        \hat{\mu}_n(\data_n))
    \\
      &
        +
        \empmeas{[\psi_{\tau}(\blank; \hat{a}_{1,n}(\data_n), \hat{a}_{2,n}(\data_n),
        \hat{b}_n(\data_n), \hat{\pi}_n(\data_n))]}
    \\
    = &
        \empmeas{[\psi_{\tau}^1(\blank; \hat{a}_{1,n}(\data_n), \hat{a}_{2,n}(\data_n),
        \hat{b}_n(\data_n), \hat{\pi}_n(\data_n))]}
    \\
      &
        - \empmeas{[\psi_{\tau}^0(\blank; \hat{a}_{1,n}(\data_n), \hat{a}_{2,n}(\data_n),
        \hat{b}_n(\data_n), \hat{\pi}_n(\data_n))]}.
  \end{split}
\end{equation}

@@

In this section we report results from a simulation study without competing
risks. In the next section, we consider a setting with competing risks. Here we
first compare the state learner to an IPCW based discrete super learner that
uses either the Kaplan-Meier estimator or a Cox model to estimate the censoring
probability. We next compare the state learner to the super learner proposed by
\cite{westling2021inference}.
% ANDERS: kan du skrive her hvad forskellen er mellem og formaalet er
% med de to simulationsstudierscenarier?
% i den ene sammenligner vi med IPCW i den anden med westling, ik?
% Maaske en figur (flowchart) eller tabel kunne hjaelpe laeseren at faa overblik.
% Maaske separation i to subsections kunne hjaelpe yderliger
%
% THOMAS: Jeg ved ikke, om jeg ville sige at der er to simulationstudier, der er
% bare to ting vi undersøge med det samme studie. Data er genereret på samme
% måde til at undersøge de to spørgsmål. Har opdateret introen til afsnittet, er
% det bedre?

@@

We compare the state learner to two IPCW based super learners: The
first superlearner, called IPCW(Cox), uses a Cox model with main
effects to estimate the censoring probabilities, while the second
super learner, called IPCW(KM), uses the Kaplan-Meier estimator to
estimate the censoring probabilities. The Cox model for the censoring
distribution is correctly specified in both simulation settings while
the Kaplan Meier estimator only estimates the censoring model
correctly in the simulation setting where censoring is
independent. Both IPCW super learners are fitted using the
\texttt{R}-package \texttt{riskRegression}
\citep{Gerds_Ohlendorff_Ozenne_2023}.
% ANDERS: det er uklar for mig hvordan
% de super learners valger/kombinerer de learners der ingaar i library: er det
% Brier score? over time?
% 
% THOMAS: De er diskrete super learner baseret på Score function med summary
% measure 'ibs', så det er integrated brier score.
Each super learner provides a learner for the cumulative hazard
function for the outcome of interest. From the cumulative hazard
function a risk prediction model can be obtained (c.f.,
equation~(\ref{eq:surv-def}) with $\Lambda_2 = 0$). We measure the
performance of each super learner by calculating the index of
prediction accuracy (IPA) \citep{kattan2018index} at a fixed time
horizon (36 months) for the risk prediction model provided by the
super learner. The IPA is 1 minus the ratio between the model's Brier
score and the null model's Brier score, where the null model is the
model that does not use any covariate information. The IPA is
approximated using a large (\( n = 20,000 \)) independent data set of
uncensored data. As a benchmark we calculate the performance of the
risk prediction model chosen by the oracle selector, which uses the
large data set of uncensored event times to select the learner with
the highest IPA.

@@

\begin{figure}
  \centering %
  \includeFigCond[width=1\linewidth]{experiment-fig-sl-ipcw.pdf}
  \caption[]{For the risk prediction models provided by each of the super
    learners, the IPA is plotted against sample size. The results are averages across 
    1000 simulated data sets and the error bars are used to quantify the Monte Carlo
    uncertainty.
    % ANDERS: er det korrekt at det er averages across 1000 data sets?
    %
    % THOMAS: Ja. ...?
  }
  \label{fig:ipcw-fail}
\end{figure}

@@

The results are shown in Figure~\ref{fig:zelefski}. We see that for
most sample sizes, the state learner selected prediction models for
both censoring and outcome which have similar or higher IPA compared
to the prediction models selected by \texttt{survSL}.
% ANDERS: it is a bit lazy to write "for most sample sizes" :) it is
% slightly annoying that the y-axis scale changes in the 4
% figures. would that be better with Brier instead of IPA?
%
% THOMAS: No, it is worse with Brier score I seem to recall. I have fixed to
% y-axis horizontally now, it works well for the outcome model but poorly for
% the censoring. Let me know what you like the best.

\begin{figure}
  \centering %
  \includeFigCond[width=1\linewidth]{experiment-fig-sl-survSL.pdf}
  \caption[]{For the risk prediction models of both the outcome and the
    censoring probability provided by each of the super learners, the IPA is
    plotted against sample size. The results are averages across 1000 repetitions and
    the error bars are used to quantify the Monte Carlo uncertainty.}
  % ANDERS: would be good to know how how many trainings set individuals were
  % right censored before the evaluation time point of the IPA. maybe of
  % interest how you chose the evaluation time point.
  %
  % THOMAS: Good point, but I think that the time of evaluation did not matter
  % too much. Of course, if you pick a very early time point where nothing has
  % happened, there is almost no difference in performance.
  \label{fig:zelefski}
\end{figure}

@@

% ANDERS: har brug for en gentle intro saetning som fortaeller hvad der kommer:
% in this section we re-analyse the ... Det skal ogsaa vaere klart i starten af
% sektion af der kommer noget om average treatment effect.
%
% THOMAS: Yes, klart. Fungerer det her:
In this section we re-analyze the prostate cancer data analyzed by
\cite{kattan2000pretreatment}, which we introduced in
Section~\ref{sec:numer-exper}. The data consists of 1,042 patients and includes
a competing event in the form of death without tumor recurrence. We use the
state learner to rank learners of the cause-specific cumulative hazard
functions. We then use the highest ranked combination of learners and the
estimation strategy outlined in Section~\ref{sec:targeted-learning} to estimate
the average treatment effect of hormone therapy on death and tumor recurrence.

We consider death without tumor recurrence and recurrence of tumor as competing
risks. We include the fives learner \texttt{KM}, \texttt{Cox strata CS},
\texttt{Lasso}, \texttt{Elastic}, and \texttt{RF} which are described in
Table~\ref{tab:zel-library}. We use the same library of learners to learn
\( \Lambda_1 \), \( \Lambda_2 \), and $\Gamma$. In this case, \( \Lambda_1 \)
denotes the cause-specific cumulative hazard function of tumor recurrence, and
\( \Lambda_2 \) denotes the cause-specific cumulative hazard function of death
without tumor recurrence.

@@

We use the learners of the three cumulative hazard functions
selected by the state learner and the estimator defined in
Section~\ref{sec:cause-spec-aver}, equation~(\ref{eq:one-step-comp-ate}), to
estimate the average treatment effect of hormone therapy on risk of tumor
recurrence and death. The propensity score is estimated with a lasso model that
includes all levels of interaction. The results are shown in
Figure~\ref{fig:zelefski-real-target} for 6 month intervals after baseline with
point wise 95\% confidence intervals. We see that hormone therapy appears to
decrease the risk of tumor recurrence and increase the risk of death without
tumor recurrence, but that none of the estimated effects are statistically
significant from zero.

@@

We have proposed a new super learner that can be used with
right-censored data and competing events. Compared to existing
IPCW-based methods, the advantage of the state learner is that it does
not depend on a pre-specified estimator of the censoring distribution,
but selects one automatically based on a library of learners for the
censoring distribution. Furthermore, the state learner neither
requires that \( \Lambda_j \) can be written as an integral with respect to
Lebesgue measure nor assumes a
% ANDERS: "density of \Lambda_j" er uklar hvordan er denne density defineret?
%
% THOMAS: Better now?
(semi-)parametric form for $\Lambda_j$. In the remainder of this
section we point to some limitations of our proposal and discuss
avenues for further research.

A major advantage of the state learner is that the performance of each
combination of learners can be estimated without (semi)parametric
assumptions or additional nuisance parameters. The drawback with this
approach is that we are rarely interested in features of the observed
data distribution when the data are right-censored: we are only
interested in the uncensored event time distribution. The finite
sample oracle inequality in Corollary~\ref{cor:oracle-prop} concerns
the function \( F \), which is a feature of \( P \in \mathcal{P} \),
while what we are typically interested in is \( \Lambda_j \) or
\( S \), which are features of \( Q \in \mathcal{Q} \). We emphasise
that while the state learner provides us with estimates of
\( \Lambda_j \) and $\Gamma$ based on libraries \( \mathcal{A}_j \)
and \( \mathcal{B} \), the performance of these learners is not
assessed directly for their respective target parameters, but only
indirectly via the performance of \( F \).
% ANDERS: er du enig med min omformulering af den sidste saetning?
%
% THOMAS: Yes, super godt. 
For settings without competing risks, our numerical studies suggest
that measuring the performance of \( F \) also leads to good
performance for estimation of \( S \). However, further research on
this issue, both numerical and theoretical, is warranted.
% ANDERS: ved ikke helt hvad du taenker man kan goer i fremtiden men synes
% egentig at den sidste saetning kunne godt slettes. :)
%
% THOMAS: Forstår ikke helt din kommentar. Hvilken sætning, den ovenover din
% kommentar eller den nedenunder? :)

@@

Our proposed super learner can be implemented with a broad library of learners
and using existing software.
% For instance the \texttt{R}-package
% \texttt{riskRegression} \citep{Gerds_Ohlendorff_Ozenne_2023} and the \texttt{R}-package by westling
Furthermore, while
the library \( \mathcal{F}(\mathcal{A}_1,\mathcal{A}_2,\mathcal{B}) \) consists
of \( |\mathcal{A}_1||\mathcal{A}_2||\mathcal{B}| \) many learners, we only need to fit
\( |\mathcal{A}_1| +|\mathcal{A}_2| + |\mathcal{B}| \) many learners in each fold. To
evaluate the performance of each learner we need to perform
\( |\mathcal{A}_1||\mathcal{A}_2||\mathcal{B}| \) many operations to calculate the
integrated Brier score in each hold-out sample, one for each combination of the
fitted models, but these operations are often negligible compared to fitting the
models. Hence the state learner is essentially not more computationally demanding
than any procedure that uses super learning to learn $\Lambda_1$, $\Lambda_2$,
and $\Gamma$ separately. While our proposal is based on constructing the library
\( \mathcal{F} \) from libraries for learning \( \Lambda_1 \), $\Lambda_2$, and
$\Gamma$, it would also be of interest to consider learners that estimate
\( F \) directly.
% ANDERS: how would one estimate F? using HAL? or RF? but such a learner would
% not provide risk prediction models or would it? if yes you could say it
%
% THOMAS: I don't think that estimators for F exists -- that is what Vi is
% working on, right? Given F we can always obtain \Lambda_1 and \Lambda_2
% (equation 8 or eq:7 in tex) and then we can get predictions from these.
